---
title: "Predicting type of Iris Flower"
author: "Ravi Teja, RiteshDev, Bharath, Yeshwanth"
date: "19/11/2020"
output:
  html_document: default
  pdf_document: default
---

## Loading Packages

```{r setup, include=FALSE}
library(caret)
```

## Load the data


```{r iris}
data("iris")
dataset <- iris
```

## Creating Validation

You can also embed plots, for example:

```{r validation, echo=FALSE}
validation_index <- createDataPartition(dataset$Species, p = 0.80, list = FALSE)
validation <- dataset[-validation_index,]
dataset <- dataset[validation_index,]
```

## Summarizing the data

### Dimensions of Data Set

``` {r dimensions}
dim(dataset)
```

### Types of Attributes

```{r types}
sapply(dataset, class)
```

### Sample Look at Data

```{r samplelook}
head(dataset)
```

### Levels of class

```{r levels}
levels(dataset$Species)
```

### Class Distribution

```{r classdistribution}
percentage <- prop.table(table(dataset$Species))*100
cbind(freq = table(dataset$Species),percentage = percentage)
```

# Statistical Summary

```{r Statistical Summary}
summary(dataset)
```

## Visualization

### Univariate Plots

```{r univariateplots}
x <- dataset[, 1:4]
y <- dataset[,5]
par(mfrow=c(1,4))
for (i in 1:4) {
  boxplot(x[,i], main = names(iris)[i])
}
plot(y)
```

### MultiVariate Plots

```{r multivariateplots}
featurePlot(x=x, y=y, plot = "ellipse")
featurePlot(x=x, y=y, plot = "box")
scales <- list(x = list(relation = "free"), y = list(relation = "free"))
featurePlot(x=x, y=y, plot = "density", scales = scales)
```

## Testing Different Algorithms


### X-Fold Cross Validation

```{r crossvalidation}
control <- trainControl(method = "cv", number = 10)
metric <- "Accuracy"
```

### Build Models

Let’s evaluate 5 different algorithms:

#### Linear Discriminant Analysis (LDA)
#### Classification and Regression Trees (CART).
#### k-Nearest Neighbors (kNN).
#### Support Vector Machines (SVM) with a linear kernel.
#### Random Forest (RF)



```{r models}
set.seed(7)
fit.lda <- train(Species~., data=dataset, method="lda", metric=metric, trControl=control)

set.seed(7)
fit.cart <- train(Species~., data=dataset, method="rpart", metric=metric, trControl=control)

set.seed(7)
fit.knn <- train(Species~., data=dataset, method="knn", metric=metric, trControl=control)

set.seed(7)
fit.svm <- train(Species~., data=dataset, method="svmRadial", metric=metric, trControl=control)

set.seed(7)
fit.rf <- train(Species~., data=dataset, method="rf", metric=metric, trControl=control)

```

### Select the Best Model

```{r bestmodel}
results <- resamples(list(lda=fit.lda, cart=fit.cart, knn=fit.knn, svm=fit.svm, rf=fit.rf))
summary(results)
dotplot(results)
```

The results for just LDA model can be summarized

```{r lda}
print(fit.lda)
```

## Making Predictions

```{r predictions}
predictions <- predict(fit.lda, validation)
confusionMatrix(predictions, validation$Species)
predictions
predictions2 <- predict(fit.rf, validation)
confusionMatrix(predictions2, validation$Species)
acc.out <- confusionMatrix(predictions2, validation$Species)$overall["Accuracy"]
```
**`r acc.out`**
Despite the small validation dataset but this result is in the expected margin 97% +/- 4% suggesting we have an accurate and reliable accurate model