<h3 position = "right">Statistical Inference Project</h3>
by, <h4><b><i>Ravi Teja L</i></b></h4>
*22-06-2020*
</br>
<h2 style = "background-color :#AFAFAF; color: black"><center>Part-1</center></h2>
</br>
<h3><center><b>
Comparing Exponential Distribution in R to Central Limit Theorem</b></center></h3>
</br>
### **Instructions**

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

## Loading Libraries

```{r load , results = 'hide'}
library("data.table")
library("ggplot2")
library("dplyr")
library(tidyr)
library(ggpubr)
```

## Production

```{r prod}
# set seed
set.seed(31)

# set lambda
lambda <- 0.2

# samples
n <- 40

# simulations
simulations <- 1000

samples <- matrix(sample(rexp(n, lambda), n*simulations, replace = TRUE), n, simulations)

sampleMeans <- apply(samples, 2, mean)
```

# **Question-1**

Show the sample mean and compare it to the theoretical mean of the distribution.

```{r}
obtainedMean <- mean(sampleMeans)
obtainedMean
basicMean <- 1/lambda
basicMean
t.test(sampleMeans)$conf
t.test(sampleMeans)$p.value
```

* Based on the comparision we can understand the fact that the true mean must be definitely greater than 5 i.e., Theoretical Mean.
* This is due to the fact that the theoretical mean is less than the 95% confidence interval of resampled data.
* Hence we reject the Null Hypothesis that True Mean is equal to 5.

```{r visualize}
hist(sampleMeans, xlab = "Means", ylab = "Simulations", density = 30, col = "grey")
abline(v = basicMean, col = "red")
abline(v = obtainedMean, col = "blue", )
text(x = obtainedMean,y = 0,"Sample Mean")
```

# **Question-2**

Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution

```{r variability}
sampleSds <- sd(sampleMeans)
sampleSds
basicSd <- (1/lambda)/sqrt(n)
basicSd
sampleVariances <- sampleSds^2
sampleVariances
basicVariance <- basicSd^2
basicVariance
```

* Standard Deviation of the distribution is 0.805329 with the theoretical SD calculated as 0.7905694. The Theoretical variance is 0.625. 
* The actual variance of the distribution is 0.6485548

# **Question-3**
Show that the distribution is approximately normal

```{r normal}
ggplot(data.frame(sampleMeans), aes(x=sampleMeans)) +
geom_histogram(aes(y = ..density..), binwidth=.25, colour="black", fill="grey")+
stat_function(fun = dnorm, lwd = 2, col = 'red', args = list(mean = mean(sampleMeans), sd = sd(sampleMeans))) 
```

From the above graph we can tell that the data is normally distributed.

```{r}
ggqqplot(sampleMeans)
  
```

As all the points fall approximately along this reference line, Hence we can assume normality.
Due to Due to the **Central limit theorem (CLT)**, the distribution of averages of 40 exponentials is very close to a normal distribution.

</br>
<h2 style = "background-color :#AFAFAF; color: black"><center>Part-2</center></h2>
</br>
<h3><center><b>
Comparing Exponential Distribution in R to Central Limit Theorem</b></center></h3>
</br>

## Instructions

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.

# Load the ToothGrowth data and perform some basic exploratory data analyses

```{r load}
data("ToothGrowth")
```

```{r modify}
ToothGrowth <- data.table(ToothGrowth)
plot(x = ToothGrowth$dose, y = ToothGrowth$len, xlab = "Dose", ylab = "Lengths", col = ToothGrowth$supp)
legend(0.5, 34, legend=unique(ToothGrowth$supp), col = c("red", "black"), lty=1:2, cex=0.8)
```

From the graph, it can be observed that as the dosage increases tooth length increases.But it also appears that OJ tooth delivers more tooth growth even at lower dosages

# Basic summary of Data

```{r}
summary(ToothGrowth)
head(ToothGrowth)
```

Examining Dose and Supplement
```{r}
ToothGrowth[, .(meanlen = mean(len), Count = .N), by = .(dose,supp)]
```

# Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose


By the summary the main relation camparision is between the usages of supplement to tooth lengths.
So the tests are to be made on supplements and tooth length for different type of dosages
```{r}
lowdose <- ToothGrowth[ToothGrowth[, dose == 0.5]]
middose <- ToothGrowth[ToothGrowth[, dose == 1.0]]
highdose <- ToothGrowth[ToothGrowth[, dose == 2.0]]
```

```{r}
t.test(ToothGrowth$len ~ ToothGrowth$supp)
```

The above test does show a wide confidence interval but the significance of p-value shows higher value so we can conclude and donot reject Null Hypothesis.

```{r}
t.test(lowdose$len ~ lowdose$supp, paired = FALSE)
```

From the above test we can find a lot of variation between means of two supplements and low p-value.
That definitely concludes the fact that OJ shows better results that VC

```{r}
t.test(middose$len ~ middose$supp, paired = FALSE)
```

From the above test we can find a lot of variation between means of two supplements and low p-value.
That definitely concludes the fact that OJ shows better results that VC

```{r}
t.test(highdose$len ~ highdose$supp, paired = FALSE)
```

From the above test we can find very low variation between two supplements.
Here we can conclude that OJ and VC show similar results in higher dosages

# Conclusions and Assumptions

Based off this data we can conclude the following:

As the dosage increases, tooth length increases regardless of what kind of supplement we use.
At the 0.5 mg and 1.0 mg dosage the OJ supplement method provides more tooth growth than that of VC Supplement
At the 2.0 mg dosage, there is no significant difference between the OJ and VC supplement methods.

**Assumptions**

We assume that the measurements are not paired.
We assume the populations are independent, that there was no crossover between the subjects and dosage.

On an overall scenario we can consider that OJ performs better than VC in most of the cases


